---
title: "stats506hw3"
author: "leon lin"
format: html
---

[GitHub Repository](https://github.com/the-rizzard-of-0z/stats506.git) 

```{r}
setwd("/Users/leonlin/Documents/stats506local/stats506")
library(haven)
library(tidyverse)
library(knitr)
library(DBI)
library(RSQLite)
library(stringr)
```

1. 
a. 
```{r}
DEMO_D <- read_xpt("data/DEMO_D.XPT")
VIX_D <- read_xpt("data/VIX_D.XPT")
merged_data <- inner_join(DEMO_D, VIX_D, by = "SEQN")
cat("Total sample size after merging is:", nrow(merged_data), "\n")
```

b.
```{r}
table_1_b <- merged_data %>% 
  mutate(Age_Bracket = cut(RIDAGEYR, breaks = seq(0, 80, by = 10), include.lowest = TRUE,
                           labels = c("0-9", "10-19", "20-29", "30-39", "40-49", 
                                      "50-59", "60-69", "70-79"))) %>% 
  group_by(Age_Bracket) %>%
  filter(!(is.na((Age_Bracket))) & !(is.na((VIQ220)))) %>%
  summarize(Proportion = round(sum(VIQ220 == 1) / n(), 3)) 

# Output table.
table_1_b %>%
  kable(
    format = "html", 
    col.names = c("Age Bracket", "Proportion"),
    caption = "Proportion of Respondents Wearing Glasses/Contact Lenses for Distance Vision by Age Group",
    align = c("c")
  ) 
```

c. 
Note our refrence levels for gender: male, and race/ethnicity: Mexican american
```{r}
# Ignore missing values or "Dont know" values. Rename factor variables. 
merged_data <- merged_data %>%
  filter(VIQ220 == 1 | VIQ220 == 2) %>%
  mutate(
    VIQ220 = ifelse(VIQ220 == 2, 0, VIQ220),
    # Rename factor variables with meaningful labels
    RIAGENDR = factor(RIAGENDR, levels = c(1, 2), labels = c("Male", "Female")),
    RIDRETH1 = factor(RIDRETH1, levels = c(1, 2, 3, 4, 5), 
                      labels = c("Mexican American", "Other Hispanic", 
                                 "Non-Hispanic White", "Non-Hispanic Black", 
                                 "Other Race"))
  )
  
  
# Fit models.
model1 <- glm(VIQ220 ~ RIDAGEYR, family = binomial, data = merged_data)
model2 <- glm(VIQ220 ~ RIDAGEYR + factor(RIDRETH1) + factor(RIAGENDR), family = binomial, data = merged_data)
model3 <- glm(VIQ220 ~ RIDAGEYR + factor(RIDRETH1) + factor(RIAGENDR) + INDFMPIR, family = binomial, data = merged_data)
```

```{r}
#' Extract Model Odds Ratios
#' 
#' @param model a fitted logistic reg model
#' @return odds ratios of model coeffs
get_odds_ratios <- function(model) {
  odds_ratios <- exp(coef(summary(model))[, "Estimate"])
  return(odds_ratios)
}

#' Calculate Pseudo R^2
#' 
#' @param model a fitted logistic reg model
#' @return pseudo r^2
pseudo_r2 <- function(model) {
  1 - (model$deviance / model$null.deviance)
}

#' creates table presenting the estimated odds ratios,  sample size for the model, pseudo-R2, and AIC value for a model. 
#' 
#' @param model a fitted logistic reg model
#' @param name desired name of the model
#' @return table depicting
create_table <- function(model, name) {
  odds_ratios <- round(get_odds_ratios(model),3)
  
  odds_df <- as.data.frame(t(odds_ratios))
  colnames(odds_df) <- names(odds_ratios)
  
  # clean col names
  colnames(odds_df) <- gsub("factor\\(|\\)|RIDRETH1|RIAGENDR", "", colnames(odds_df))
  colnames(odds_df) <- gsub("\\(|\\)", "", colnames(odds_df))
  colnames(odds_df)[colnames(odds_df) == "RIDAGEYR"] <- "Age"
  colnames(odds_df)[colnames(odds_df) == "INDFMPIR"] <- "PIR"
  
  odds_df$Model <- name
  odds_df$Sample_Size <- model$nobs 
  odds_df$Pseudo_R2 <- round(pseudo_r2(model), 3)
  odds_df$AIC <- round(AIC(model), 3)
  
  return(odds_df)
}
```

```{r}
model1_table <- create_table(model1, "Model 1")
model2_table <- create_table(model2, "Model 2")
model3_table <- create_table(model3, "Model 3")

summary_table <- bind_rows(model1_table, model2_table, model3_table) %>% 
  select(Model, everything())

summary_table %>%
  kable(
    format = "html", 
    caption = "Model Summary Table: Estimated Odds Ratios, Sample Size, Pseudo-R², and AIC",
    align = c("c")
  ) 
```

d. 
Recall that in our logistic regressions, our reference level for gender is: Male.

Proceed with Wald's test to test whether the odds of men and women being wears of glasess/contact lenses for distance vision differs while controlling for race, age, and PIR.
H0: coeff_female_model3 = 1 Ha: coeff_female_model3 != 1
```{r}
coef_gender_female <- summary(model3)$coefficients["factor(RIAGENDR)Female", "Estimate"]
se_gender_female <- summary(model3)$coefficients["factor(RIAGENDR)Female", "Std. Error"]

# Walds statsitc
wald_stat <- coef_gender_female / se_gender_female
p_value_wald <- 2 * (1 - pnorm(abs(wald_stat)))

paste0("p-value: ", p_value_wald)
```

Since p-value small, there is sufficient evidence to reject the null hypothesis. There is evidence to support the claim that the odds of wearing glasses/contact lenses for distance vision differ between men and women. 

Proceed with Fisher's Exact test to test whether the proportion of wearers of glasses/contact lenses for distance vision differs between men and women.
H0: proportion male - proportion female = 0, Ha:proportion male - proportion female != 0
```{r}
contingency_table <- table(merged_data$RIAGENDR, merged_data$VIQ220)
fisher_test_result <- fisher.test(contingency_table)
fisher_test_result
```
Since p-value small, there is sufficient evidence to reject the null hypothesis. There is evidence to support the claim that the proportion of wearers of glasses/contact lenses for distance vision differs between men and women.

2. 
```{r}
sakila <- dbConnect(RSQLite::SQLite(), "data/sakila_master.db")
```

a.
```{r}
dbGetQuery(sakila, "
SELECT release_year, COUNT(*) as movie_count
    FROM film
  GROUP BY release_year
  ORDER BY release_year ASC
  LIMIT 1;
")
```
Thus, the oldest movie came from 2006, and 1000 movies were released that year.

b. 
```{r}
film <- dbGetQuery(sakila,"SELECT * FROM film")
category <- dbGetQuery(sakila,"SELECT * FROM category")
film_category <- dbGetQuery(sakila,"SELECT * FROM film_category")

film_category %>% 
  left_join(category, by = "category_id") %>%
  group_by(name) %>% 
  summarize(count = n()) %>% 
  arrange(count) %>% 
  head(5)

dbGetQuery(sakila, "
SELECT c.name, COUNT(fc.film_id) AS count
    FROM film_category fc
  JOIN category c ON fc.category_id = c.category_id
  GROUP BY c.name
  ORDER BY count ASC
  LIMIT 5;
")
```
We see that the genre least common in the data is music, with 51 movies. 

c. 
```{r}
country <- dbGetQuery(sakila,"SELECT * FROM country")
city <- dbGetQuery(sakila,"SELECT * FROM city")
address <- dbGetQuery(sakila,"SELECT * FROM address")
customer <- dbGetQuery(sakila,"SELECT * FROM customer")

customer %>% 
  left_join(address, by = "address_id") %>%
  left_join(city, by = "city_id") %>% 
  left_join(country, by = "country_id") %>%
  group_by(country) %>%
  summarize(total_customers = n()) %>% 
  filter(total_customers == 13)

dbGetQuery(sakila, "
SELECT country, COUNT(*) AS total_customers
    FROM 
    customer LEFT JOIN address ON customer.address_id = address.address_id
      LEFT JOIN city ON address.city_id = city.city_id
        LEFT JOIN country ON city.country_id = country.country_id
GROUP BY country
HAVING COUNT(*) = 13;
")
```
We see that Argentina and Nigeria have exactly 13 total customers. 

3.
```{r}
us_500 <- read.csv("data/us-500.csv")
```

a.
```{r}
us_500 %>%
  mutate(email_tld = str_extract(email, "\\.[a-z]+$")) %>%
  summarize(proportion = mean(email_tld == ".com"))
```
b.
```{r}
us_500 %>%
  mutate(has_non_alphanumeric = str_detect(email, "[^a-zA-Z0-9@.]")) %>%
  summarize(proportion = mean(has_non_alphanumeric))
```

c. 
```{r}
us_500 %>%
  mutate(area_code1 = str_extract(phone1, "^\\d{3}"), 
         area_code2 = str_extract(phone2, "^\\d{3}")) %>%
  pivot_longer(cols = c(area_code1, area_code2), names_to = "type", values_to = "area_code") %>%
  group_by(area_code) %>% 
  summarise(n=n()) %>% 
  arrange(-n) %>% 
  head(6)
```
Hence, the most 5 popular area codes are 973, 212, 215, 410, and 201.

d. 
```{r}
us_500 %>%
  mutate(apartment_number = as.numeric(str_extract(address, "\\d+$"))) %>%
  filter(!is.na(apartment_number)) %>%
  mutate(log_apartment_number = log(apartment_number)) %>%
  ggplot(aes(x = log_apartment_number)) +
  geom_histogram(binwidth = 2, fill = "skyblue", color = "black", boundary = 0) +  # Set bin boundary to 0
  labs(title = "Histogram of Log of Apartment Numbers",
       x = "Log of Apartment Number",
       y = "Frequency") +
  theme_minimal()

```

```{r}
# extract numeric vector of leading digits
leading_digits <- us_500 %>%
  mutate(apartment_number = as.numeric(str_extract(address, "\\d+$"))) %>%
  filter(!is.na(apartment_number)) %>% 
  mutate(leading_digit = substr(as.character(apartment_number), 1, 1)) %>%
  pull(leading_digit) %>% 
  as.numeric()

# we conduct a chi-sq test to see if the distribution of leading digits from the data follows benford's law.
benford_probs <- log10(1 + 1 / (1:9))
observed <- table(factor(leading_digits, levels = 1:9))
chisq.test(observed, p = benford_probs)
```
Since p-value is small, there is sufficient evidence to reject the null hypothesis that the distribution of the observed leading digits of apartment numbers follows Benford's Law.






